---
title: "world money_ball - kaggle dataset"
author: "patrick mcdevitt (& collaboration with preeti swaminathan)"
date: "04 juin 2017"
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(ggplot2)
library(MASS)
library(car)
library(moments)
library(cowplot)
library(grid)
library(gridExtra)

home_dir <- "~/_smu/_src/money_ball/"
setwd(home_dir)

data_dir <- "./data"

```

## money_ball - Regression Project -


### __Intro__  

*** 
The following __Context__ and __Content__ is from : https://www.kaggle.com/unsdsn/world-money_ball  



#### __Objective__

Within the context of __World money_ball Day__ explore the data available from the the United Nations survey and idneitfy the characteristics that are associated with higher __money_ball__ scores

*** 
### __Descriptive statistics__  

*** 

### __Analysis__  

*** 

### __Interpretation__  

*** 

### __Conclusion__  

*** 


  
  
#### __Setup some data directories  

```{r read data, include = FALSE, message = FALSE}

	setwd(home_dir)
	setwd(data_dir)

	mlb <- read.csv("moneyball.csv", stringsAsFactors = FALSE)
	setwd(home_dir)

	names(mlb) <- tolower(names(mlb))
	
	for (i in 2:(length(mlb)))
	{
		if (class(mlb[,i]) == "character")
		{
			mlb[,i] <- factor (mlb[,i])
		}
	}
```

#### __Remove extreme outliers from each column of numerical data__  

```{r derived variable, echo = FALSE, message = FALSE, comment = "", fig.width = 12}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	create derived variable - number of singles
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	mlb$team_batting_1b <- mlb$team_batting_h -
							mlb$team_batting_2b - 
							mlb$team_batting_3b - 
							mlb$team_batting_hr

# ...	Slugging percentage is the number of total bases divided by the number of at bats.
# ...	Its formula is ([Singles] + [Doubles x 2] + [Triples x 3] + [Home Runs x 4])/[At Bats]

	mlb$team_batting_ab <- mlb$team_batting_h + (27 * 162)

	mlb$team_batting_slg <- mlb$team_batting_1b + 
							2 * mlb$team_batting_2b +
							3 * mlb$team_batting_3b +
							4 * mlb$team_batting_hr / mlb$team_batting_ab

# ...	OBP = (Hits + Walks + Hit by Pitch) / (At Bats + Walks + Hit by Pitch + Sac Flies)
	
	mlb$team_batting_obp <- rowSums(mlb[,c("team_batting_h", "team_batting_bb", "team_batting_hbp")],
							na.rm = TRUE)
	
	mlb$team_batting_obp <- mlb$team_batting_obp /
							(rowSums(mlb[,c("team_batting_ab", "team_batting_bb", "team_batting_hbp")],
							na.rm = TRUE) + 0.3 * 162)
	
```


```{r impute missing values to median, include = FALSE, message = FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	impute missing values to median
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	for (i in 3 : length(mlb))
	{
		if(class(mlb[,i]) == "integer" || class(mlb[,i]) == "numeric")
		{
			mlb[,i][is.na(mlb[,i])] = median(mlb[,i], na.rm = TRUE)
		}
	}

```



```{r ln transform on left skewed variables, include = FALSE, message = FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	ln transform on left skewed variables
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	for (i in 3 : length(mlb))
	{
		if(class(mlb[,i]) == "integer" || class(mlb[,i]) == "numeric")
		{
			if (skewness(mlb[,i], na.rm = TRUE) >= 0.9)
			{
				mlb[,i] <- log(mlb[,i] + 1)
			}
		}
	}

```


```{r remove 5sigma outliers, include = FALSE, message = FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	remove outliers ... more than 5 sigma from mean value
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	for (i in 3 : length(mlb))
	{
		if(class(mlb[,i]) == "integer" || class(mlb[,i]) == "numeric")
		{
			mlb[,i][which(scale(mlb[,i]) > 5)] <- NA
			mlb[,i][which(scale(mlb[,i]) < -5)] <- NA
		}
	}

# ...	remvoe rows where wins percentage is outside actual experience
# ...		according to wikipedia
# ...			least wins in modern era = 0.235 percentage
# ...			most wins in modern era = 0.763 percentage

	mlb <- mlb[!(mlb$target_wins / 162) < 0.235,]
	mlb <- mlb[!(mlb$target_wins / 162) > 0.763,]

```


```{r plot base data, echo = FALSE, message = FALSE, comment = "", fig.width = 12}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	make some plots for numeric variables .....
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	cx_scale = 3
# ... 	use cx_scale = 1 for in-line html knitted docs
# ...	use cx_scale = 3 for pdf landscape mode knitted docs

	par (mfrow = c (1, 2))

	plot(mlb$target_wins, col = "darkgoldenrod",
		cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
	  	xlab = names(mlb[i]),
	  	main = "Dependent variable - Target Wins sequence")

	plot(mlb$target_wins/162, col = "sienna3",
		cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
	  	xlab = names(mlb[i]),
	  	main = "Dependent variable - Percentage Wins")

	for (i in 3 : length(mlb))
	{
		if(class(mlb[,i]) == "integer" || class(mlb[,i]) == "numeric" || class(mlb[,i]) == "matrix")
		{
				  
#			png (filename = paste0(names(mlb[i]), "_%02d.png"), width = 1800, height = 600)
			par (mfrow = c (1, 4))

			plot(mlb[,i], col = "cyan",
				 cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
				  	xlab = names(mlb[i]),
				  	main = "Independent variable sequence")

			hist(mlb[,i], main = names(mlb[i]), cex.lab = cx_scale, cex.main = cx_scale)
			if (skewness(mlb[,i], na.rm = TRUE) < 0)
			{
				txt_pos <- "topleft"
			}
			else {
				txt_pos <- "topright"
			}
			legend(txt_pos, legend = c(paste("Mean =", round(mean(mlb[,i], na.rm = TRUE), 1)),
                            paste("Median =",round(median(mlb[,i], na.rm = TRUE), 1)),
                            paste("Std.Dev =", round(sd(mlb[,i], na.rm = TRUE), 1)), 
                            paste("Skewness =", round(skewness(mlb[,i], na.rm = TRUE), 1))), 
                  bty = "n", cex = cx_scale)
			
			
			plot(mlb$target_wins  ~ mlb[,i], col = "blue", cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
				 	ylab = "money_ball Score",
				  	xlab = names(mlb[i]),
				  	main = "Independent variable vs. money_ball Score")
			abline(lm(mlb$target_wins ~ mlb[,i]))

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	look at residuals from one-variable linear fit			
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

			fit <- lm(mlb$target_wins ~ mlb[,i], na.action = na.exclude)
			res <- resid(fit)
			plot (mlb[,i], res, cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
				  ylab = "Residuals",
				  xlab = names(mlb[i]),
				  main = "Residuals from Linear Fit",
				  col = "magenta")
			abline (0, 0)
			
		}
#		dev.off()
	}

```


#### __Make first data visualization plot for each column of dataset__  


```{r baseline lm, echo = FALSE, message = FALSE, comment = "", fig.width = 12}


	model_df <- mlb
	model_df$index <- NULL
#	model_df$team_batting_hbp <- NULL
#	model_df$team_batting_1b <- NULL
#	model_df$team_batting_h <- NULL
#	model_df$team_batting_tb <- NULL
#	model_df$team_baserun_cs <- NULL

	first_model <- lm(target_wins ~ ., data = model_df, na.action = na.exclude)
	first_pred <- predict(first_model)
	plot (model_df$target_wins, first_pred,
		  cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
		  ylab = "Observed Wins",
		  xlab = "Predicted Wins",
		  main = "Pred vs. Observed from Linear Fit",
		  col = "magenta")
	
	summary(first_model)
```

```{r full lm with interactions, echo = FALSE, message = FALSE, comment = "", fig.width = 12}
#	model_df$team_batting_1b <- NULL

	model_df <- mlb
	model_df$index <- NULL
#	model_df$team_batting_hbp <- NULL
#	model_df$team_batting_1b <- NULL
#	model_df$team_batting_h <- NULL
#	model_df$team_batting_tb <- NULL
#	model_df$team_baserun_cs <- NULL

	interxn_model <- lm(target_wins ~ .*., data = model_df, na.action = na.exclude)
	interxn_pred <- predict(interxn_model)
	plot (model_df$target_wins, interxn_pred,
		  cex = cx_scale, cex.lab = cx_scale, cex.axis = cx_scale, cex.main = cx_scale,
		  ylab = "Observed Wins",
		  xlab = "Predicted Wins",
		  main = "Pred vs. Observed from Linear Fit with Interactions",
		  col = "magenta")
	
	summary(interxn_model)
```



```{r save subsetted data, include = TRUE, message = FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	save data frame for SAS input file
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	moneyball_sas_input <- mlb

	sas_dir <- "~/sas/SASUniversityEdition/myfolders/stats_ii"
	setwd(sas_dir)
	write.csv (moneyball_sas_input, file = "moneyball_training_set_cleaned.csv", row.names = FALSE, na = ".")
	
	setwd(home_dir)
	setwd(data_dir)
	write.csv (moneyball_sas_input, file = "moneyball_training_set_cleaned.csv", row.names = FALSE, na = ".")
	setwd(home_dir)
	

```
